---
title: "Reproducible Research"
output: html_document
---



**Loading and preprocessing the data.**

```{r, echo = FALSE}
setwd("C://Users/Ane4ka/Desktop/Oleg/Reproducible Research")
```
```{r}
activity = read.csv("activity.csv")
activity$date = as.Date(activity$date)
```

**What is mean total number of steps taken per day?**

For this part of the assignment, you can ignore the missing values in the dataset.

```{r}
data1 = na.omit(activity)
```

Calculate the total number of steps taken per day

```{r}
totalStepsPerDay = tapply(data1$steps, data1$date, sum)
```

Make a histogram of the total number of steps taken each day

```{r}
barplot(totalStepsPerDay)
```

Calculate and report the mean and median of the total number of steps taken per day

```{r}
meanWithMiss = mean(totalStepsPerDay)
meanWithMiss
medianWithMiss = median(totalStepsPerDay)
medianWithMiss
```

We can see that the median `r medianWithMiss` is a little different from the mean `r round(meanWithMiss, digits = 0)`.

**What is the average daily activity pattern?**

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
avgStepInterval = tapply(data1$steps, data1$interval, mean)
Interval = as.integer(names(avgStepInterval))
plot(Interval, avgStepInterval, type = 'l')
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
intervalMaxStep = as.integer(names(avgStepInterval[which(avgStepInterval == max(avgStepInterval))]))
intervalMaxStep
```

We can see that `r intervalMaxStep` 5-minute interval contains the maximum number of steps.

**Imputing missing values**

Number of days/intervals where there are missing values

```{r}
data.na = activity[!complete.cases(activity),]
length(unique(data.na$date))
length(unique(data.na$interval))
```

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
nrow(data.na)
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy is the mean for that 5-minute interval.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
data2 = activity
for (i in 1:nrow(data2)) {
  if (is.na(data2$steps[i])) {
    for (j in 1:length(avgStepInterval)) {
      if (as.integer(names(avgStepInterval)[j]) == data2$interval[i]) {
        data2$steps[i] = avgStepInterval[j]
      }
    }
  }
}
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
totalStepsPerDay2 = tapply(data2$steps, data2$date, sum)
barplot(totalStepsPerDay2)
meanWithoutMiss = mean(totalStepsPerDay2)
meanWithoutMiss
medianWithoutMiss = median(totalStepsPerDay2)
medianWithoutMiss
```

We can see that when we use strategy for filling in all of the missing values in the dataset as the mean for that 5-minute interval, the median `r medianWithoutMiss` and the mean `r meanWithoutMiss` are equal. Alse the means with strategy `r meanWithoutMiss` and without one `r meanWithMiss` are equal, but the median with strategy `r medianWithoutMiss` and without one `r medianWithMiss` are a little different.

**Are there differences in activity patterns between weekdays and weekends?**

Use the dataset with the filled-in missing values for this part.

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
newVar = vector(mode = "character", length = nrow(data2))
for (i in 1:nrow(data2)) {
  if (weekdays(data2$date[i]) == "воскресенье" | weekdays(data2$date[i]) == "суббота") {
    newVar[i] = "weekend"
  } else {
    newVar[i] = "weekday"
  }
}
data2$newVar = as.factor(newVar)
```


Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
subWeekend = subset(data2, newVar == "weekend")
subWeekday = subset(data2, newVar == "weekday")

avgStepWeekend = tapply(subWeekend$steps, subWeekend$interval, mean)
intervalWeekend = as.integer(names(avgStepWeekend))
plot(intervalWeekend, avgStepWeekend, type = 'l')

avgStepWeekday = tapply(subWeekday$steps, subWeekday$interval, mean)
intervalWeekday = as.integer(names(avgStepWeekday))
plot(intervalWeekday, avgStepWeekday, type = 'l')
```

